---
title: "Introduction to covidBR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to covidBR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  comment = "#>"
)
```

The `covidBR` package provides easy access to official COVID-19 data sets of Brazil. All functions' syntax is based on a simple logic that allows users to download updated data sets for all years. This vignette provides an overview of `covidBR`.

### Installation 

You can install the development version of `covidBR` from [GitHub](https://github.com/) with:

```{r eval = FALSE, include = TRUE}
# Development version
# install.packages("devtools")
devtools::install_github("arianacabral/covidBR")
```

### Load package

Now, let's load the `covidBR` and the libraries we’ll use in this vignette.

```{r setup, warning=FALSE, message=FALSE, error=FALSE}
# Load packages
library(covidBR)
library(fs)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(knitr)
```

### General usage

The `covidBR` allows you to obtain the *URL* on the [Covid-19 Portal](covid.saude.gov.br) available for downloading the files. 

> Daily, the Ministry of Health (**Ministério da Saúde**) updates and publicly disseminates the database, and, consequently, the *download URL* is updated.

```{r, error=FALSE, message=FALSE, warning=FALSE}
# Get updated URL to download the public file
download_url <- covidBR::cbr_download_url()

# Print URL
download_url
```

Now let's download the public file available at the obtained URL.
```{r, eval=FALSE, include=TRUE}
# Download public file
resp <- covidBR::cbr_download(url_file = download_url,
                              keep_filename = TRUE)

```

A `zip` file has been downloaded from the [COVID-19 Portal](https://covid.saude.gov.br/). We use the `cbr_unzip()` function to unzip it.

```{r eval=FALSE, include=TRUE}
# Get filename
filename <- fs::path_file(resp$url)

# Print filename
filename

# Unzip
covidBR::cbr_unzip(zip_file = filename,
                   keep_zip_file = FALSE) # Delete the compressed file
```

Simply use the `cbr_data()` function to combine all files and generate a single dataset.

```{r eval=FALSE,include=FALSE}
# Bind multiple files
covidBR.db <- covidBR::cbr_data()
```

### Available variables

You can view what variables are available using the `str()` function.

```{r eval=FALSE, include=TRUE}
# Display the dataset's structure
str(covidBR.db)
```

You can also view the data in a tabular form.

```{r eval=FALSE, include=TRUE}
covidBR.db %>% 
  dplyr::filter(data >= (Sys.Date()-2), # last 2 days
                municipio== "") %>% # only Brazilian states
  dplyr::select(-c(municipio, coduf, 
                   codmun, cod_regiao_saude,
                   nome_regiao_saude,
                   semana_epi,
                   populacao_tcu2019,
                   interior_metropolitana,
                   em_acompanhamento_novos, 
                   recuperadosnovos)) %>% 
  knitr:: kable(digits = 4,
                caption = "Brazil (last 2 days)")
```
