---
title: "Introduction to covidBR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to covidBR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  comment = "#>"
)
```

The `covidBR` package provides easy access to official COVID-19 data sets of Brazil. All functions' syntax is based on a simple logic that allows users to download updated data sets for all years. This vignette provides an overview of `covidBR`.

### Installation 

You can install the development version of `covidBR` from [GitHub](https://github.com/) with:

```{r eval = FALSE, include = TRUE}
# Development version
# install.packages("devtools")
devtools::install_github("arianacabral/covidBR")
```


### Load packages

Now, let's load the `covidBR` and the libraries we’ll use in this vignette.

```{r setup, warning=FALSE, message=FALSE, error=FALSE}
# Load packages
library(covidBR)
library(fs)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(knitr)
```

### General usage

The `covidBR` allows you to obtain the *URL* on the [Covid-19 Portal](covid.saude.gov.br) available for downloading the files. 

> Daily, the Ministry of Health (**Ministério da Saúde**) updates and publicly disseminates the database, and, consequently, the *download URL* is updated.

```{r, error = FALSE, message = FALSE, warning = FALSE}
# Get updated URL to download the public file
download_url <- covidBR::cbr_download_url()

# Print URL
download_url
```

Now let's download the public file available at the obtained URL.

```{r eval=FALSE, include=TRUE}
# Download public file
resp <- covidBR::cbr_download(url_file = download_url,
                              dir = here::here("inst/extdata"), #saving to "extradata" folder
                              keep_filename = TRUE)

# Get filename
filename <- fs::path_file(resp$url)

# Print filename
filename

```

A `zip` file has been downloaded from the [COVID-19 Portal](https://covid.saude.gov.br/). We use the `cbr_unzip()` function to unzip it.

```{r eval=FALSE, include=TRUE}
# Set absolute directory
abs_path <- stringr::str_c(
  here::here("inst/extdata"),
  filename,
  sep = "/"
)

# Unzip
covidBR::cbr_unzip(zip_file = abs_path,
                   dir = here::here("inst/extdata"), # the directory to extract archive contents
                   keep_zip_file = FALSE) # Delete the compressed file
```

Simply use the `cbr_data()` function to combine all files and generate a single dataset.

> You can already combine the files by specifying a time range of interest. 
For example, use the argument `date = Sys.Date()` to retrieve today's data.

```{r message=FALSE, error=FALSE, warning=FALSE}
# Bind multiple files
covidBR.db <- covidBR::cbr_data(dir = here::here("inst/extdata"))  
```

### Available variables

You can view what variables are available using the `str()` function.

```{r message=FALSE, error=FALSE, warning=FALSE}
# Display the dataset's structure
str(covidBR.db)
```

### Data exploration

You may be interested in specific information, and we will now demonstrate how to work with some case scenarios.

#### Data exploration by date

For this purpose, we can use the package `lubridate`. It is an R package that makes it easier to work with dates and times.

> See `help(package = "lubridate")`.

How can I get this year's data?

```{r eval=TRUE, include=TRUE}
# Get current date
today <- Sys.Date()

# Print current date
today

# Get this year
current_year <- lubridate::year(today)

# Print the current year 
current_year

# Set first day of this year
day_1 <- lubridate::ymd(paste0(current_year,"01","01",sep="-"))

# Print the first day of this year 
day_1

# Get this year's data
covidBR.db <- covidBR::cbr_data(dir = here::here("./inst/extdata"),
                                date = c(day_1, 
                                         today)) 

# Tabular view
covidBR.db %>% 
  dplyr::select(regiao, 
                data,
                casos_acumulado,
                casos_novos, 
                obitos_novos) %>% 
  head(n = 10L) %>% 
  knitr::kable() 
```

How can I get the last six months' worth of data?

```{r eval=TRUE, include=TRUE}
# Get current date
today <- Sys.Date()

# Print current date
today

# Get date from 6 months ago
date_6_months_ago <- today - months(6)

# Print date from 6 months ago
date_6_months_ago

# Get data from the last 6 months
covidBR.db <- covidBR::cbr_data(dir = here::here("./inst/extdata"),
                                date = c(date_6_months_ago,today)) 

# Tabular view
covidBR.db %>% 
  dplyr::select(regiao, 
                data,
                casos_acumulado,
                casos_novos, 
                obitos_novos) %>% 
  head(n = 10L) %>% 
  knitr::kable()

```

How can I get the last 30 days' worth of data?

```{r eval=TRUE, include=TRUE}
# Get current date
today <- Sys.Date()

# Print current date
today

# Get date from 30 days ago
date_30_days_ago <- today - lubridate::days(30)

# Print date from 30 days ago
date_30_days_ago

# Get data from the last 30 days
covidBR.db <- covidBR::cbr_data(dir = here::here("./inst/extdata"),
                                date = c(date_30_days_ago, today)) 

# Tabular view
covidBR.db %>% 
  dplyr::select(regiao, 
                data,
                casos_acumulado,
                casos_novos, 
                obitos_novos) %>% 
  head(n = 10L) %>% 
  knitr::kable()
```

#### Data visualization

```{r}
# Obtain data for Brazil
tidycovidBR <- covidBR::cbr_data(
  dir = here::here("inst/extdata"))  %>% 
  dplyr::filter(regiao == "Brasil") 

# Select the interesting columns 
tidycovidBR <- tidycovidBR %>% 
  dplyr::select(c(regiao, data,
                  casos_acumulado:obitos_novos))

# Tabular view 
tidycovidBR %>% 
  head(n = 15L) %>% 
  knitr::kable()

# Summary
summary(tidycovidBR)

# Check for duplicate dates
is_unique <- length(unique(tidycovidBR$data)) == length(tidycovidBR$data)

# Plot data
if(is_unique){
  
  tidycovidBR %>% 
  dplyr::ungroup() %>% 
  tidyr::pivot_longer(
    cols = -c(regiao,data),
    names_to  = "variavel",
    values_to = "valor") %>% 
  dplyr::mutate_at("variavel", 
                   stringr::str_replace_all,
                   pattern = "_",
                   replacement = " ")  %>% 
  ggplot(aes(x = data, y = valor)) + 
  geom_line() + 
  facet_wrap(~ variavel, scales = "free_y", ncol = 1) + 
  labs(x = "Data da notificação", 
       y = "") + 
  theme_minimal() 
  
}

```
