---
title: "Introduction to covidBR dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to covidBR dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  comment = "#>"
)
```

The `covid` R package includes a dataset containing **COVID-19** epidemiological data from Brazil. In this introductory vignette, we will highlight some of the characteristics of this dataset that make it useful for reporting COVID-19 statistics in Brazil.

## The covidBR package

The original data comes from the Brazilian COVID-19 Portal (**Portal do COVID-19**), and can be found at <https://covid.saude.gov.br>. Daily, the Ministry of Health (**Ministerio da Saúde**) updates and publicly disseminates the database. In this context, the `covidBR` R package was built to extract data from the Brazilian COVID-19 Portal.  

This package contains a dataset called `covidBR`.
Furthermore, it provides raw data, downloaded from the COVID-19 Portal, in `CSV` format files (access [here](https://github.com/arianacabral/covidBR/tree/master/data-raw)). 

### Installation 

You can install the development version of `covidBR` from [GitHub](https://github.com/) with:

```{r eval = FALSE, include = TRUE}
# Development version
# install.packages("devtools")
devtools::install_github("arianacabral/covidBR")
```

### Load packages

Now, let's load the `covidBR` and the libraries we will use in this vignette.

```{r setup, warning=FALSE, message=FALSE, error=FALSE}
# Load packages
library(covidBR)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(rmarkdown)
```

## Meet the covidBR dataset

The `covidBR::covidBR` dataset contains 17 variables for each Brazilian region, municipality, and state.

> You can read more about the variables by typing `?covidBR::covidBR`.

You can view what variables are available using the `str()` function.

```{r error=FALSE,message=FALSE,warning=FALSE}
# Display the dataset's structure
str(covidBR)
```

You can also view the data in a tabular form.

```{r}
covidBR %>%
  dplyr::filter(data >= max(data),
                municipio== "") %>%
  dplyr::select(!c(municipio, coduf,
                   codmun, cod_regiao_saude,
                   nome_regiao_saude,
                   semana_epi,
                   populacao_tcu2019,
                   interior_metropolitana,
                   em_acompanhamento_novos,
                   recuperadosnovos)) %>%
  dplyr::arrange(regiao, estado) %>%
  rmarkdown::paged_table(options = list(rows.print = 49))
```

> Note that the raw data contains duplicate information!
