---
title: "Introduction to covidBR dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to covidBR dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  comment = "#>"
)
```

The `covid` R package includes a dataset containing **COVID-19** epidemiological data from Brazil. In this introductory vignette, we will highlight some of the characteristics of this dataset that make it useful for reporting COVID-19 statistics in Brazil.

## The covidBR package

The original data comes from the Brazilian COVID-19 Portal (**Portal do COVID-19**), and can be found at <https://covid.saude.gov.br>. Daily, the Ministry of Health (**Ministerio da Saúde**) updates and publicly disseminates the database. In this context, the `covidBR` R package was built to extract data from the Brazilian COVID-19 Portal.  

This package contains a dataset called `covidBR`.
Furthermore, it provides raw data, downloaded from the COVID-19 Portal, in `CSV` format files (access [here](https://github.com/arianacabral/covidBR/tree/master/data-raw)). 

### Installation 

You can install the development version of `covidBR` from [GitHub](https://github.com/) with:

```{r eval = FALSE, include = TRUE}
# Development version
# install.packages("devtools")
devtools::install_github("arianacabral/covidBR")
```

### Load packages

Now, let's load the `covidBR` and the libraries we will use in this vignette.

```{r setup, warning=FALSE, message=FALSE, error=FALSE}
# Load packages
library(covidBR)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(knitr)
```

## Meet the covidBR dataset

The `covidBR::covidBR` dataset contains 17 variables for each Brazilian region, municipality, and state.

> You can read more about the variables by typing `?covidBR::covidBR`.

You can view what variables are available using the `str()` function.

```{r error=FALSE,message=FALSE,warning=FALSE}
# Display the dataset's structure
str(covidBR)
```

You can also view the data in a tabular form.

```{r}
covidBR %>%
  dplyr::filter(data >= max(data),
                municipio== "") %>%
  dplyr::select(-c(municipio, coduf,
                   codmun, cod_regiao_saude,
                   nome_regiao_saude,
                   semana_epi,
                   populacao_tcu2019,
                   interior_metropolitana,
                   em_acompanhamento_novos,
                   recuperadosnovos)) %>%
  dplyr::arrange(regiao, estado) %>% 
  knitr::kable()
```

> Note that the raw data contains duplicate information!

### Tidy data for more efficient data science 

In this case, it is necessary to organize data to help us work in an efficient, reproducible, and collaborative way. 

The example below shows one way to organize the data and remove duplicate elements, but you can represent this same data in several ways.

```{r error=FALSE,message=FALSE,warning=FALSE}
# Get covidBR dataset
tidycovidBR <- covidBR

# Retain all rows with the most recent notification date for each Brazilian state.
tidycovidBR <- tidycovidBR %>% 
  dplyr::filter(data >= max(data),
                municipio == "") %>%
  dplyr::arrange(regiao, estado)  

# Pivot the offending columns into a new pair of variables
tidycovidBR <- tidycovidBR %>% tidyr::pivot_longer(
  cols = c(populacao_tcu2019:interior_metropolitana,codmun),
  names_to  = "variavel",
  values_to = "valor")

# Select the interesting columns
tidycovidBR <- tidycovidBR %>% 
  dplyr::select(-c(cod_regiao_saude,nome_regiao_saude))

# Aggregate the duplicate columns
tidycovidBR <- tidycovidBR %>% 
  dplyr::group_by(variavel,data,regiao,estado,municipio,
                  semana_epi, coduf) %>% 
  dplyr::summarise_at(vars(valor),
                      list(valor = sum), 
                      na.rm = TRUE)
# View the data in the original structure
tidycovidBR %>% 
  tidyr::pivot_wider(names_from = variavel, 
                     values_from = valor) %>%
  dplyr::ungroup() %>%  
  dplyr::select(c(regiao,estado,data,
                  casos_acumulado,casos_novos,
                  obitos_acumulado, obitos_novos)) %>% 
  dplyr::arrange(regiao, estado, data) %>% 
  knitr::kable()

```

### Visualization

You can now use [`ggplot2`](https://ggplot2.tidyverse.org/index.html) to explore the data and create visually appealing plots. See an example.

```{r error=FALSE,message=FALSE,warning=FALSE}
tidycovidBR  %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(variavel != "codmun" &
                  variavel != "interior_metropolitana" &
                  variavel != "interior_metropolitana" & 
                  variavel != "populacao_tcu2019" &
                  variavel != "recuperadosnovos", 
                regiao != "Brasil"
  ) %>%  
  dplyr::mutate_at("variavel", 
                   stringr::str_replace_all,
                   pattern = "_",
                   replacement = " ")  %>% 
  ggplot(aes(x = estado, y = valor, fill = estado)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ variavel, scales = "free_y", ncol = 1) + 
  labs(x = "", y = "") + 
  theme_minimal() + 
  theme(legend.position = "none")
```

